\chapter{Device Lending in PCI Express Networks}
\label{paper:NOSSDAV}
\paperthumb

\begin{description}
	\item[Authors:]
		Lars Bj{\o}rlykke Kristian, \textbf{Jonas Markussen}, H{\aa}kon Kvale Stensland,
		Michael Riegler, Hugo Kohmann, Friedrich Seifert, Roy Nordstr{\o}m, Carsten Griwodz, P{\aa}l Halvorsen

	\item[Abstract:]
		The challenge of scaling \lgls{io}{IO} performance of multimedia systems to demands
		of their users has attracted much research.
		A lot of effort has gone into
		development of distributed systems that add little latency and computing overhead.
		For machines in \lgls{pcie}{PCI Express~(PCIe)} clusters,
		we propose Device Lending as a novel solution which works at a system
		level.
		%
		Device Lending achieves low latency and extremely low computing overhead without
		requiring \textit{any} application-specific distribution mechanisms.
		For the application, the remote \lgls{io}{IO} resource appears local.
		In fact, even the drivers of the operating system remain unaware that
		hardware resources are located in remote machines.
		%
		By enabling machines in a \lgls{pcie}{PCIe} cluster to lend a wide variety of hardware, 
		cluster machines can get temporary access to a pool of \lgls{io}{IO} resources. 
		Network cards, \lgls{fpga}{FPGAs}, \lgls{ssd}{SSDs}, and even \lgls{gpu}{GPUs} can easily 
		be shared among computers.
		Our proposed solution, Device Lending, works transparently without requiring any modifications to drivers,
		operating systems or software applications.

	\item[Candidate's contributions:]
		Markussen was the main author of the paper, responsible for writing most of the text and organizing the collaboration with 
		all of the authors. He designed and performed the performance evaluation of the Device~Lending method described in the paper.
		Markussen had several discussions with Kristiansen about Kristiansen's original implementation of Device~Lending, 
		and contributed to its development through testing and evaluating performance benchmarks.

	\item[Published in:]
		Proceedings of the 26th International Workshop on Network and Operating Systems Support for Digital Audio and Video~(NOSSDAV'16),
		May~2016, article~10, pp.~10:1--10:6.
		\doi{10.1145/2910642.2910650}

	\item[Contributed to:]
		TODO

\end{description}
\includearticle[]{papers/NOSSDAV-2016}
