\chapter{Acknowledgements}
I'd like to express my sincerest gratitude to my friend and colleague, Lars Bj{\o}rlykke Kristiansen.
%
This dissertation would simply not have been possible without your support and help. 
%
Thank you for sharing my frustration over confusing lab results, the many hours spent teaching me about technical details, as well as your countless insights, technical understanding, and visionary ideas for things to try out and experiments to do.
%
I'm looking forward to working with you for many years to come.


I'd also like to thank to my PhD advisors, H{\aa}kon Kvale Stensland, Carsten Griwodz, and P{\aa}l Halvorsen.
%
Not only have you been extremely patient with me, more than anyone can reasonably expect from any advisor, but your invaluable feedback and input has truly elevated the quality of my work.
%
Your innovative approach to computer science and dedication to high quality research inspires me, and I hope we may collaborate on future research projects.
%
I also want to thank all of my colleagues at Dolphin, especially my bosses Hugo Kohmann and Roy Nordstr{\o}m, who have allowed me time and resources to perform this work---even when it didn't align perfectly with tasks that were more urgent for the company.


Thanks to my family and supportive friends, particularly Katrine, Anders, Christian, Martin, Axel, Morten, Thomas, Karoline, Nora, Jonatan, and Kristian.
%
Thank you for continuing to believe in me, even when I doubted myself. 
%
It is because of your continued support I didn't give up after all these years.



I'd also like to thank my friends and former colleagues at Simula: Michael, Bendik, Halvor, Konstantin, Kristoffer R., and Preben.
%
You made the working environment of Simula educational and also very fun for me.
%
I have many fun memories from conference trips abroad, general office humor, and playing pool and joking around.
%
Finally, I would like to thank the Simula Premium Coffee Club (aka. ``The Secret Society of Aromatic Snobbery'') for providing the excellent coffee consumed at many late hours as deadlines were approaching, thus saving me from the awful, black substance produced by the (so-called) ``coffee'' machine.


\vskip\onelineskip
\begin{flushleft}
    \sffamily
    \uiocolon\textbf{\theauthor}
    \\
    Oslo,\MONTH\the\year
\end{flushleft}
